Polymethylation Scores for Prenatal Maternal Smoke Exposure Persist Until Age

### 2021-09-30 Reports/Poster/IGES_poster_vertical.Rmd
poster changes

```{r}
outcome_labels=c('Global methylation', 'Pediatric clock', 'Polymethylation score \n (newborn, cell-type controlled, mean centered)')
names(outcome_labels)=c('globalmethylation', 'pediatric', 'SSnewbornCT_center')
modeldata%>%dplyr::select(idnum, childteen, SSnewbornCT_center, globalmethylation, pediatric)%>%pivot_longer(cols=c(globalmethylation, pediatric, SSnewbornCT_center), names_to='Methylation summary measure type', values_to='Methylation summary measure')%>%pivot_wider(names_from=childteen, values_from=`Methylation summary measure`)%>%ggplot(aes(x=C, y=T))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+facet_wrap(~`Methylation summary measure type`, scales='free', labeller = labeller(`Methylation summary measure type`=outcome_labels))
```

```{r}
modeldata %>% pivot_longer(cols=c('globalmethylation', 'pediatric', 'SSnewbornCT_center'), names_to='measuretype', values_to='Methylation summary measure')%>%ggplot(aes(x=childteen, y=`Methylation summary measure`, group=idnum))+geom_point()+geom_line(alpha=0.1, color='grey')+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~measuretype, scales='free_y', labeller = labeller(measuretype=outcome_labels))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))
```

@@

data_text=local_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.01)%>%
  unnest(tidied)%>%
  filter(outcome=='Global methylation' & term=='smkPreg_binaryYes'& predictors=='secondhand smoke exposure model')

local_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='secondhand smoke exposure model')%>%
  filter(outcome%in% my_outcomes)%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome, scales='free_x', labeller = labeller(outcome=label_wrap_gen(15)))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.5)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_color_discrete('', limits=c('Age 9', 'Age 15'))+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry'))

>>>>

```{r}
outcome_labels=c('Global methylation', 'Pediatric clock', 'Polymethylation score \n (newborn, mean centered)')
names(outcome_labels)=c('globalmethylation', 'pediatric', 'SSnewbornCT_center')

modeldata%>%dplyr::select(idnum, childteen, SSnewbornCT_center, globalmethylation, pediatric)%>%pivot_longer(cols=c(globalmethylation, pediatric, SSnewbornCT_center), names_to='Methylation summary measure type', values_to='Methylation summary measure')%>%pivot_wider(names_from=childteen, values_from=`Methylation summary measure`)%>%mutate(meth_f=factor(`Methylation summary measure type`, levels=c('SSnewbornCT_center', 'globalmethylation', 'pediatric')))%>%ggplot(aes(x=C, y=T))+geom_point()+stat_cor()+geom_smooth(method=lm)+theme_classic()+scale_x_continuous('Age 9 score')+scale_y_continuous('Age 15 score')+facet_wrap(~meth_f, scales='free', labeller = labeller(meth_f=outcome_labels))+theme(text=element_text(size=16))

```{r}
modeldata %>%group_by(idnum)%>%filter(n()>1)%>%pivot_longer(cols=c('globalmethylation', 'pediatric', 'SSnewbornCT_center'), names_to='measuretype', values_to='Methylation summary measure')%>%group_by(measuretype, idnum) %>%
    mutate(slope = (`Methylation summary measure`[childteen=='T'] - `Methylation summary measure`[childteen=='C'])/(2-1))%>%filter(!is.na(`Methylation summary measure`))%>%
  ggplot(aes(x=childteen, y=`Methylation summary measure`, group=idnum))+geom_point()+geom_line(alpha=0.1, aes(color=slope>0))+geom_smooth(aes(group=1), method='lm', fill='blue')+facet_wrap(~measuretype, scales='free_y', labeller = labeller(measuretype=outcome_labels))+theme_classic()+scale_x_discrete('', labels=c('Age 9', 'Age 15'))+scale_color_manual(name='', labels=c('slope<0', 'slope>0'), values=c('blue', 'red'))

modeldata %>%group_by(idnum)%>%filter(n()>1)%>%pivot_longer(cols=c('globalmethylation', 'pediatric', 'SSnewbornCT_center'), names_to='measuretype', values_to='Methylation summary measure')%>%group_by(measuretype, idnum) %>%
    mutate(slope = (`Methylation summary measure`[childteen=='T'] - `Methylation summary measure`[childteen=='C'])/(2-1), slope_factor=case_when(slope>0~'Positive', slope<=0~'Negative'))%>%filter(!is.na(`Methylation summary measure`))%>%ggplot(aes(x=slope))+geom_histogram()+geom_blank(aes(x=-slope)) + facet_wrap(~measuretype, scales='free', ncol=1)
```

@@

all_models=rbind(local_models, global_models%>%mutate(ancestry='Unstratified'))

data_text=all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n), x=-0.25)%>%
  unnest(tidied)%>%
  filter(outcome=='Sustained smoking polymethylation score (newborns, cell-type controlled)' & term=='smkPreg_binaryYes'& predictors=='secondhand smoke exposure model')%>%mutate(outcome_f=outcome)

names(outcome_labels)=my_outcomes
all_models%>%
  mutate(n=map_dbl(data, nrow), n=paste0('n=', n))%>%
  unnest(tidied)%>%
  filter(term=='smkPreg_binaryYes' & predictors=='secondhand smoke exposure model')%>%
  filter(outcome%in% my_outcomes)%>%
    mutate(outcome_f=factor(outcome, levels=rev(my_outcomes)))%>%
  ggplot(aes(x=estimate, xmin=conf.low, xmax=conf.high, color=factor(age, rev(levels(age))), y=ancestry))+
  geom_point(position = position_dodge(width = 1))+
  geom_errorbarh(height=0.1, position = position_dodge(width = 1))+
  facet_grid(.~outcome_f, scales='free_x', labeller = labeller(outcome_f=outcome_labels))+
  geom_vline(xintercept=0, color='red', linetype="dashed", alpha=0.5)+
  geom_text(data=data_text, aes(x=x,label=n), position=position_dodge(width = 1), hjust=0.05)+
  theme_classic()+
  theme(text = element_text(size=16))+theme(legend.position = 'bottom')+
  scale_color_discrete('', limits=c('Age 9', 'Age 15'))+
  scale_x_continuous('Coefficient for prenatal maternal smoking + 95% CI')+
  scale_y_discrete('', limits=c('European ancestry', 'Hispanic ancestry', 'African ancestry', 'Unstratified'))+scale_color_manual('', values=c('black', 'darkgrey'))
